---
title: "McCollum_Breck_HW2"
output: html_document
---
Homework #2 preliminary warm up stuff

```{r}
library(gapminder)
library(ggplot2)
library(dplyr)

head(gapminder)
```

Okay, now that you recall what Gapminder looks like, begin trying to figure out how to make plot 1. Start with the general structure, then get fancier with shapes, colors, regressions, trend lines etc...

```{r}
ggplot(data = gapminder, aes(x = gdpPercap, y = lifeExp)) + geom_point()

ggplot(data = gapminder, aes(x = gdpPercap, y = lifeExp, color = continent)) + geom_point()

ggplot(data = gapminder, aes(x = gdpPercap, y = lifeExp, shape = continent)) + geom_point()

ggplot(data = gapminder, aes(x = gdpPercap, y = lifeExp, color = continent, shape = continent)) + geom_point()
```

Whoohoo! that didn't take too long to figure out! I had a feeling that it wouldn't let me assign two different "aesthics" to continent, but it didn't actually fight me at all!  



####Now on to transformations...

"ggplot(data = gapminder, aes(x = gdpPercap, y = lifeExp, color = continent, shape = continent)) + geom_point()" was the line of code that got us to this point, but as much fun as it is to copy and paste that where ever I need it, I think I ought to give it a code name. How about "p1" for plot 1?  

```{r}
p1 <- ggplot(data = gapminder, aes(x = gdpPercap, y = lifeExp, color = continent, shape = continent)) + geom_point()
```

### Now lets see how we can transform these data...

?log()

### Some moderatley helpful info there, but I believe I only want to transform the x values. I am going to do a little trial and error here, since I don't have a good grasp on transformations in any capacity

ggplot(data = gapminder, aes(x = log(gdpPercap), y = lifeExp, color = continent, shape = continent)) + geom_point()

### Ooooh that is super interesting! I apparently chose a different transformation than you did, according to our differing x axis, but I think this still counts as a 'transformation', hopefully you won't dock points because our axes are different. (I was trying to make it look like yours, I just have no idea what kind of transformation you used).

### Now I better re-nickname this since I only transformed a part of p1, I'll stick with convention and call this one "p2"


p2 <- ggplot(data = gapminder, aes(x = log(gdpPercap), y = lifeExp, color = continent, shape = continent)) + geom_point()

?geom_line()

### Hmmmm, that wasn't helpful. I am trying google. Google gave me an interesting thing to test out. Perhaps I want to include a line of best fit rather than just a linear line. These guys, http://docs.ggplot2.org/current/geom_abline.html say I can calculate intercept and slope for my data using "coef(lm(y ~ x, data = my data set))" where I fill in my "y, x, and data set". Lets give it a shot...

coef(lm(lifeExp ~ gdpPercap, data = gapminder))

### Super cool! I am going to run with this for a bit but with what that function reported: (Intercept)    gdpPercap  5.395556e+01 7.648826e-04

### Next the docs.ggplot2.org guys say I can type the intercept and slope values into a geom deal called "geom_abline", like this:
### "geom_abline(intercept = 37, slope = -5)" and add it to my p2.

p2 + geom_abline(intercept = 5.395556e+01, slope = 7.648826e-04)

### Alright, that is very interesting, but apparently a line of best fit is not what I want here. Still, "geom_abline" might be useful in the future. Cool.

### I keep seeing things online about "geom_smooth" but I don't really understand what that is. Let's try it...

p2 + geom_smooth(model = lm)

### Okay, that made it unhappy. Let's try something else...

p2 + geom_smooth(method = lm)

### Whoa crazy, but not what I am looking for...I need more info about 'geom_smooth'...

?geom_smooth

### Nope still doesn't make sense. Back to google...

p2 + geom_smooth(model = "lm", se = TRUE)
p2 + geom_smooth(method = "x", se = TRUE)
p2 + geom_line()

### Yikes, undo! Now I am headed in the wrong direction...There must be some way I can do an overall linear regression, rather than one for each continent...

ggplot(data = gapminder, aes(x = log(gdpPercap), y = lifeExp, color = continent, shape = continent)) + geom_point() + geom_smooth(method = lm)

ggplot(data = gapminder, aes(x = log(gdpPercap), y = lifeExp, color = continent, shape = continent)) + geom_point() + geom_smooth(method = log(gdpPercap))

ggplot(data = gapminder, aes(x = log(gdpPercap), y = lifeExp, color = continent, shape = continent)) + geom_point() + geom_smooth(x = log(gdpPercap))

###Fuuuuuuuuuuuuuck. I don't know about this one. This (from http://www.r-bloggers.com/how-to-draw-the-line-with-ggplot2/) looks a bit complicated but I'll try it... 

### plot.y <- qplot(y=Temp, x=Wind, data=data)
### model.y <- lm(Temp ~ Wind, data)
### coef.y <- coef(model.y)
### plot.y + geom_abline(intercept=coef.y[1],
### slope=coef.y[2])

ggplot(data = gapminder, aes(x = log(gdpPercap), y = lifeExp, color = continent, shape = continent)) + geom_point() + geom_smooth(method = lm(lifeExp ~ log(gdpPercap, gapminder))

### Nope. Try again.

model.y <- lm(lifeExp ~ log(gdpPercap, gapminder))

### Okay this is really frustrating. I am moving on before I throw my laptop out the window.


### For Plot 4, I need to histogram life expectancy for each continent by year, in a handful of tiles. This sounds really, really difficult...

ggplot(data = gapminder, aes(x = lifeExp, y = count, color = continent)) + geom_density(kernel = "gaussian")

### This gave me a weird error where it thinks x= a string of numbers (life expectancies, perhaps) and says "arguments imply differing number of rows"

head(gapminder)

ggplot(data = gapminder, aes(x = year, y = lifeExp, color = continent)) + geom_density(kernel = "gaussian")

### Hmmmm, example from docs.ggplot2.org doesn't include a y-value, maybe that is not necessary

ggplot(data = gapminder, aes(x = lifeExp, color = continent)) + geom_density(kernel = "gaussian")

### This is close, all the life expectancies are histogramed together in one plot. Now I need to separate by year, and fill in the curves with transparent colors  

ggplot(data = gapminder, aes(x = lifeExp, color = continent)) + geom_density(kernel = "gaussian") + facet_grid(year ~ fl)


ggplot(data = gapminder, aes(x = lifeExp, color = continent)) + geom_density(kernel = "gaussian") + facet_grid(year ~ year)

### Hahahahaha, well that did something. 

ggplot(data = gapminder, aes(x = lifeExp, color = continent)) + geom_density(kernel = "gaussian") + facet_grid( ~ year)

### WTF? Is this correct? It looks pretty good...No, not quite. I need to put them in 3 rows rather than a line...I wonder why "fl ~ year" didn't work?

ggplot(data = gapminder, aes(x = lifeExp, color = continent)) + geom_density(kernel = "gaussian") + facet_grid(. ~ year)

ggplot(data = gapminder, aes(x = lifeExp, color = continent)) + geom_density(kernel = "gaussian") + facet_grid( ~ fl)

ggplot(data = gapminder, aes(x = lifeExp, color = continent)) + geom_density(kernel = "gaussian") + facet_wrap( ~ year)

### Okay, interesting. Maybe the ggplot2 cheat sheet is wrong on this one? Anyway, this looks good. Now, I need to fill and make fills transparent.

p4 <- ggplot(data = gapminder, aes(x = lifeExp, color = continent)) + geom_density(kernel = "gaussian") + facet_wrap( ~ year)

ggplot(data = gapminder, aes(x = lifeExp, color = continent)) + geom_density(kernel = "gaussian", fill = color) + facet_wrap( ~ year)

### Back to google...

ggplot(data = gapminder, aes(x = lifeExp, color = continent, fill= continent)) + geom_density(kernel = "gaussian") + facet_wrap( ~ year)

### Super cute, but not quite there...

ggplot(data = gapminder, aes(x = lifeExp, color = continent, fill = continent)) + geom_density(kernel = "gaussian", alpha = 0.4) + facet_wrap( ~ year)

### That looks pretty good, again thank to google for the transparency/alpha help. Not sure how to make my outlines black though... maybe color and fill are redundant?

ggplot(data = gapminder, aes(x = lifeExp, fill = continent)) + geom_density(kernel = "gaussian", alpha = 0.4) + facet_wrap( ~ year)

### HAH, word. okay I feel pretty good about that. Onward... 

### Holy mother, how much more is on this assignment?!



### Boxplot time, plot #5, this doesn't look fun either...

ggplot(data = gapminder, aes(x = continent, color = continent, fill = continent)) + geom_boxplot(lifeExp) + facet_wrap( ~ year)

### ..the fuck you mean you can't find "lifeExp"?! It's right there!!

ggplot(data = gapminder, aes(lifeExp, lifeExp, fill = continent)) + geom_boxplot() + facet_wrap( ~ year)

### Cool, but not right, huh? And I recieved 24 warnings!

warnings()

### Oh, I need continent for my x axis, somehow I ended up with lifeExp for both..

ggplot(data = gapminder, aes(continent, lifeExp, color = continent)) + geom_boxplot() + facet_wrap( ~ year)

### Hell yeah!

### Oh god, that was only Part 1, of how many Parts?! Jesus fucking christ...

### Plot what, 6 now? Where am I? I think I made this one on accident earlier...

ggplot(data = gapminder, aes(x = lifeExp, color = continent)) + geom_density(kernel = "gaussian")

### Right, just fill and make transparent I think...

ggplot(data = gapminder, aes(x = lifeExp, fill = continent)) + geom_density(kernel = "gaussian", alpha = 0.5)

### Sweeeeeeeet. I feel like that one was a gift to me, like specifically. Like, if you were here looking over my shoulder, you would say/be thinking "See, you are not a total idiot! Just wait for this next soul crushing question..."

### Plot 7 is the next soul crushing question...Okay so to find means, I think we did something like this in class recently..Too bad my notes aren't great..

### I am trying to plot mean lifeExp in Asia as histogram with a line for the mean?

ggplot(data = gapminder, aes(x = lifeExp, fill = Asia)) + geom_density(kernel = "gaussian", alpha = 0.5)

### Ugh, I hoped it would be that easy. How do I get it to select just Asia's lifeexp? Okay last week's class used "%>%" a lot, I can't for the life of me recall what that means. Perhaps ?%>% will work?

?%>%

### Nope. So, if a monkey, given an infinite amount of time hitting keys on a typewriter will eventually reproduce Hamlet (imt = infinite monkey theorem), I should be able to figure this out in 30 minutes or so, right?

gapminder %>% group_by(country) %>% slice(mean(lifeExp))

### Uh, okay. Wish I knew what that looked like..

imt <- gapminder %>% group_by(country) %>% slice(mean(lifeExp))

ggplot(data = imt, aes(x = asia, y = lifeExp, color =

ggplot(data = imt, aes(x = lifeExp, fill = continent)) + geom_density(kernel = "gaussian", alpha = 0.5)

### Okay, that is not working, lets try something else...

gapminder %>% group_by(country, lifeExp) %>% slice(mean(lifeExp))

gapminder %>% group_by(continent, lifeExp) %>% filter( %in% continent)

mean(group_by((filter(gapminder, continent == Asia)), continent)$lifeExp)
 
head(gapminder)

### Ugh, why the hell can't it find Asia??? This was the thing we typed in class, and it found 2002, with no problem!

mean(group_by((filter(gapminder, year == 2002)), continent)$pop)

mean(group_by((filter(gapminder, year == 2002)), continent)$lifeExp)

### If I ask it for mean lifeExp of 2002, it has no problem giving that to me...

mean(group_by((filter(gapminder, country == Afghanistan)), continent)$lifeExp)

str(gapminder)

### Okay, so it must treat factors differently from numbers, but how and what do I do about it?

mean(group_by((filter(gapminder[ ,c(2,4)])), continent)$lifeExp)

### Well, that is interesting. Maybe I can plot that...

imt <- mean(group_by((filter(gapminder[ ,c(2,4)])), continent)$lifeExp)

ggplot(data = imt, aes(x = lifeExp, fill = continent)) + geom_density(kernel = "gaussian", alpha = 0.5)

### Ugh, what a pain in the ass!

ggplot(data = gapminder[ ,c(2,4)], aes(x = lifeExp, fill = continent)) + geom_density(kernel = "gaussian", alpha = 0.5)

ggplot(data = gapminder[ ,c(2,4)], aes(x = lifeExp, fill = Asia)) + geom_density(kernel = "gaussian", alpha = 0.5)

ggplot(data = gapminder[ , 2], aes(x = lifeExp, fill = continent)) + geom_density(kernel = "gaussian", alpha = 0.5)


### Okay so I am not getting any closer with this one. I need to take a break and move on to the next question.

library(ggplot2)

hw_gapminder <- read.csv('./hw_gapminder.csv')

### Oh hell, yeah. I have no idea where this went, let alone how to find it...

